---
title: "AIDS Drug"
author: "Jacob Welsh"
date: "16 April 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides (triUmpsal) is under test in clinical trials. Analyze the effectiveness of this drug across the board. Consider population subgroups by gender, drug use, and sexual orientation. Finish this!

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

All questions are categorical-categorical, as a result we will use the same methods for graphical descriptions, numerical descriptions, and inferences.  
Graphical Descriptions will be Bar Charts.  
Numerical Descriptions will be cross tables.  
Inferentials wil; be a chi-square test and/or fisher test exact test.    

The conclusion for all the questions will be described at the end.  

This is true for all 5 questions.  

## Question 1 : Is the drug effective in the general population?  



$H_0$ : The drug has no effect on whether or not one gets AIDS.  

$H_A$ : The drug does have an effect on whether or not one gets AIDS.  

### Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=triUmp, type ="percent")
```

The barchart shows that the vaccine may have an effect whether one aquires aids or not. The percentage for those that got AIDS was lower than those with the placebo.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=triUmp)

rowPerc(table1)
```
This is row percentages, it shows that of the people given the drug, 19% less had contracted AIDS at 40.87, while those with the placebo were infected at 59.13 rate. It also shows that among those who did not contract AIDS, the drug and placebo made little difference.


```{r}
colPerc(table1)
```
This is a table of column percentages, it shows that of the people on the drug, 3.99 percent contracted AIDS, and 96.01 did not. 5.49 percent of people on the placebo contracted AIDS, while 94.51 percent did not contract the disease.



### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

For the chi-squared test the p-value was 0.00000002. This is less than 0.05, so we have evidence to reject the null hypothesis, implying that the drug does in fact have some sort of effect on the general population.  





## Question 2 Is the drug effective in the LGBTQ population?  


$H_0$: The drug has no effect on members in the LGBTQ population contracting AIDS.   


$H_A$: The drug does have some effect in the LGBTQ population contracting AIDS.

### Descriptive Results
```{r}
LGBTQdata = filter(triUmp, LGBTQ != "straight")

barchartGC(~infected + treatment,data=LGBTQdata , type ="percent")
```

The drug would appear to have an adverse affect on the LGBTQ community, as there is more than 60% of those who had the drug contracted AIDS.

### Numerical Results
```{r}
table2 <- xtabs(~infected + treatment, data=LGBTQdata)

rowPerc(table2)
```
This is row percentages, it shows that among people who contracted AIDS, 72.38% were on the drug, while 27.62% of them were on the placebo. This also shows that it is pretty even on those not infected,49.05% of them came from the drug group, and 50.95% came from the placebo group.  


```{r}
colPerc(table2)
```
This is a table of column percentages, it shows that of people that recieved the drug, 7.02 percent contracted AIDS, while 92.98 percent remained clean. It shows that of those on the placebo 2.7% contracted AIDS and 97.3% did not contract the disease.



### Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

Pearson's Chi-squared test has p-value = 5.029e-06, or .000005029. Because this number is smaller than .05, we will reject our null hypothesis. This means the drug DOES have an effect on the LGBTQ population, as seen above however, it is adverse and the drug increases the odds of contracting AIDS. The Fisher's test gives you the odds of being 2.7 times more likely to contract the disease.


## Question 3 Is the drug effective in the female population?  

$H_0$: The drug has no effect on whether or not females contract AIDS.  

$H_A$: The drug does have some effect on whether or not females contract AIDS?   
   
   


### Descriptive Results
```{r}
FemalesData = filter(triUmp, gender == 'F')
barchartGC(~infected + treatment,data=FemalesData, type ="percent")
```

### Numerical Results  

```{r}
table3 <- xtabs(~infected + treatment, data=FemalesData)

rowPerc(table3)
```
This is row percentages, it shows that of those who contracted AIDS 58.82% were on the test drug while the other 41.18% were on the placebo, which makes the drug look a little bad. Among those who did not contract AIDS it was once again fairly close, 48.06% were on the drug, with 51.94% being on the placebo.

```{r}
colPerc(table3)
```
This is a table of column percentages, it shows that among those on tge drug, only 3.69% contracted AIDS, while 2.42% of those on the placebo contracted the disease. The drug group escaped contracting the disease 96.31% of the time, while those on the placebo avoided it 97.58% of the time.
### Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

According the Chi-squared test the p-value = 7.686e-05, or .00007686, this small of a number means we are going to reject our null hypothesis, that this has no effect on the female population. There seems to be an effect in that females become a little bit more likely to catch the disease. The Fisher's exact test gives them a 1.5x more likely chance.


## Question 4 Is the drug effective in the male poplulation?   

$H_0$: The drug has no effect on members in the male population.  

$H_A$: The drug does have some effect in the male population.  


### Descriptive Results
```{r}
MalesData = filter(triUmp, gender == 'M')
barchartGC(~infected + treatment,data=MalesData, type ="percent")
```

At the first look of this barchart, it would seem the drug has a significant positive effect on the male population, of the males the contracted AIDS. only aroud 30% seem to have come from the group that contained the test drug.  

### Numerical Results
```{r}
table4 <- xtabs(~infected + treatment, data=MalesData)

rowPerc(table4)
```
This is row percentages, it shows that the of those who contracted AIDS, only 31.42% contracted AIDS in the group given the drug and 68.58% of the infected came from the placebo group. For the people who did not contract AIDS, as seen in all previous questions it is about the same, 50.38% were on the drug, 49.62% were not. So far it looks pretty good for the male population.  


```{r}
colPerc(table4)
```
This is a table of column percentages, it shows that of the people on the drug, 4.35% contracted AIDS, while the other 95.65% did not. It also shows that of the people on the placebo group, 9.16% contracted AIDS, while 90.84% did not. Again, looks good for the men, with the higher contraction being on the placebo side.  

### Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

With the p-value in the Chi test being somewhere around 0000000000000002.2 we are rejecting this null hypothesis, the drug most certainly has an effect on the male population in that less of them contract the disease. In the Fisher Exact test, you have odds of being half less likely to contract AIDS.  

## Question 5 Is the drug effective in the drug using community?  

$H_0$: The drug has no effect on members in the drug using community.


$H_A$: The drug does have some effect in the drug using community.

### Descriptive Results
```{r}
DrugUserData = filter(triUmp, DrugUser == "yes")
barchartGC(~infected + treatment,data=DrugUserData, type ="percent")
```

This barchart is showing some strong positive effect from the drug, here it would seem that among those who contracted AIDS it looks to be about 25% came from those on the test drug, while the other 75% came from the placebo group. Those who did not contract are again about even between groups.  


### Numerical Results
```{r}
table5 <- xtabs(~infected + treatment, data=DrugUserData)

rowPerc(table5)
```
This is row percentages, it shows that of all those who contracted AIDS, only 24.29% came from the test drug group, while 75.71% came from the placebo group, this shows a strong positive effect. Among those who did not contract AIDS 47.59% came from the group on the test drug, with the other 52.41% being from the placebo.  

```{r}
colPerc(table5)
```
This is a table of column percentages, it shows that of the people on the drug, 2.89% contracted the disease, while 97.11% did not. For the placebo group 7.77% were infected, while 92.23% did go clean. This shows again the the drug is in fact having an effect. 

### Inferential Results
```{r}
chisq.test(table5)
chisqtestGC(table5)
fisher.test(table5)
```

The Chi-squared test give us a p-value = 1.242e-07, or .0000001245. This small number means we reject the null hypothesis. There is what seems to be a strong positive effect from the test drug on the group of Drug Users, the Fisher test gives them odds at a bit more than a third of a chance to contract.  




### Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Here is a summary of the results:

Question 1: The drug is effective in preventing a person in the General Population from aquiring AIDS. It would be good to use in this population.

Question 2: The drug is effective in bringing on the contraction of AIDS in the LGBTQ community. It would not be good to use in this population.  

Question 3: The drug is effective in bringing on the contraction of AIDS in the Female population. It would not be good to use in this population.

Question 4: The drug is effective in preventing a person in the Male population from aquiring AIDS. It would be good to use in this population.

Question 5: The drug is effective in preventing a person in the Drug Using population from aquiring AIDS. It would be good to use in this population.

In three out of five of the groups, the drug perfroms to it's intended purpose of prevents AIDS. However in the LGBTQ and Female populations it is harmful and increases the likelihood of memebers of those groups to contract AIDS. This drug would likely be very beneficial for the Male and Drug using populations, but it would likely be a disaster for the Female and LGBTQ communities.